\documentclass[a4paper,12pt]{article}

% Set margins
\usepackage[hmargin=2.5cm, vmargin=3cm]{geometry}

\frenchspacing

% Language packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage[magyar]{babel}

% AMS
\usepackage{amssymb,amsmath}

% Graphic packages
\usepackage{graphicx}

% Colors
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}

% Enumeration
\usepackage{enumitem}

% Links
\usepackage{hyperref}

\begin{document}

\begin{center}
    \Large \textbf{Fuzzy Rule Approximation}
\end{center}

\section{Introduction}

The fundamental goal is to model states and their transitions.
We call \textit{state} all of the values of the considered system.
We can refer to their components as \textit{state variables}.
Therefore, the state is the set of the considered state variables.
Let denote $S$ the state which contains $n$ state variables:
\[
S = \{s_0, s_1, \ldots, s_{n-1}\}.
\]

(The explanation uses zero-based indexing, because most of the programming languages follow this convention also. It hopefully helps the implementation of the proposed method.)

The proposed behavior engine uses continuous space for the available states ($S \subseteq \mathbb{R}^n$) and it is discrete in time.
Let denote $S_0$ the initial state of the system.
The behavior engine can be denoted as a function $\mathcal{B}$:
\[S_{i+1} = \mathcal{B}(S_i).\]

From practical reasons we can categorize state variables into three different groups.
\begin{itemize}
    \item \textit{input}: Its value determined by the environment of the system.
    \item \textit{internal} or \textit{hidden}: It is necessary for and internal calculations and proper state transitions, but its value does not provide any useful information.
    \item \textit{output}: Its value considered as the result of the calculation.
\end{itemize}

From the aspect of the behavior engine this categorization is not necessary.
As a brief overview of the usage of the behavior engine can be described by the following steps.
\begin{enumerate}
    \item Define the state variables of the system.
    \item Define the relation of the state variables by rules.
    \item Set the initial values ($S_0$).
    \item Evaluate the rules and determine the next state.
    \item Read the values of the state variables (which are considered as outputs).
    \item Set the values of the state variables (which are considered as inputs).
    \item Continue from step 4.
\end{enumerate}

\section{Language Variables}

In the proposed behavior description method we can define names for the measures/amounts.
\begin{itemize}
    \item A name cannot contain spaces.
    \item The names are case sensitive.
    \item It cannot start with number.
    \item It cannot contains dot.
    \item It cannot be a keyword.
\end{itemize}
A language variable is a \{name, value\} pair.
For instance
\begin{verbatim}
  high is 30
\end{verbatim}
means that the 30 is a \emph{high} value on a particular dimension, metric, amount or measure.
(The behavior engine cannot consider its meaning.)
This expression is called \textit{statement}.

\section{Universe of Discourse}

A \textit{universe} is an ordered set of language variables, which can be referred by their name or by their index. The pairs are ordered by the values in ascending order.

Let denote $U$ a universe.
\begin{itemize}
    \item $|U|$ is the number of pairs in the set.
    \item $U[0]$ is the \emph{first} value.
    \item $U[a]$ is the value of language variable $a$.
    \item $\min(U)$ is the minimal value of $U$.
    \item $\max(U)$ is the maximal value of $U$.
\end{itemize}

For instance, let define the universe of distance in the following way.

\begin{verbatim}
In the universe of distance
  0 is close and
  10 is medium and
  50 is far.
\end{verbatim}

In this case
\[
|U| = 3, U[0] = 0, U[1] = 10, U[2] = 50, U[\text{close}] = 0, U[\text{medium}] = 10, U[\text{far}] = 50.
\]

\section{Distances on the universe}

We would like to approximate the distance between an observation and a language variable on a given universe.
We have to define a \textit{distance operator} $\delta$ for the calculation.

Let consider a universe $U$, an observation $x \in \mathbb{R}$ and a language variable called $a$.
The distance of $x$ can be calculated as
\[
\delta(U, x, a) =
\dfrac{|x - U[a])|}{\max(U) - \min(U)}.
\]
It is easy to see, that the maximal distance in the universe is 1.

For instance, for the previous example let $x = 20$. Then
\begin{align*}
&\delta(U, 20, \text{close}) = 0.4, \\
&\delta(U, 20, \text{medium}) = 0.2, \\
&\delta(U, 20, \text{high}) = 0.6. \\
\end{align*}

\section{Rules}

For illustration, let define universes for battery charge level and speed.
\begin{verbatim}
In the universe of charge_level
  0 is low and
  100 is high.

In the universe of speed
  0 is slow and
  30 is fast.
\end{verbatim}

A rule defines the relation between inputs and outputs. Therefore, a rule has two parts:
\begin{itemize}
    \item \textit{antecedent}: the input side of the rule,
    \item \textit{consequent}: the output side of the rule.
\end{itemize}

The consequent always contains only one statement, while the antecedent part can contains arbitrary number of statements joined by the \texttt{and} keyword. From this reason, we prefer to write the consequent part of the statement at the beginning of the rules.

\begin{verbatim}
The speed is slow when
  distance is medium and
  charge_level is low.

The speed is fast when
  charge_level is high.
\end{verbatim}

A rule is a point in the space of antecedent and consequent dimension. In this particular example, we are in three dimensional space ($\text{distance} \times \text{battery\_level} \times \text{speed}$). The first rule is a point in a typical sense (can be expressed by values as $(10, 0, 0)$). The second rule is a line (or by considering the bounds for this example a line segment of $(d, 100, 30), d \in [0, 50]$).

\section{Rule Distances}

For the further calculations, we have to define the distance of an observation and a rule.
Let denote its operator by $\rho$.
The distance from the rule is only depends on the antecedent part of the rule.
We can calculate the distances of the statements individually on all universes. Let denote $\delta_0, \ldots, \delta_{(k-1)}$ these distances for a rule with $k$ statements.
\[
\textbf{s} = (\delta_0, \ldots, \delta_{(k-1)}) \in [0, 1]^k,
\]
\[
\displaystyle
\rho(\textbf{s}) =
\sqrt{\dfrac{\sum_{i=0}^{k-1} \delta_i^2}{k}}
\in [0, 1].
\]
It is the normalized Euclidean distance in the $k$-dimensional space.

\section{Calculation of the Consequences}

Let denote $c_i$ the consequent value of the rules at index $i$. We would like to calculate the consequent of the rulebase.

Let denote the pairs of consequent values and the corresponding rule distances by $\textbf{r}$:
\[
\textbf{r} = \{(c_0, \rho_0), (c_1, \rho_1), \ldots, (c_{m-1}, \rho_{m-1})\}
\]
The operator $\sigma$ for consequent calculation can be defined as:
\[
\sigma(\textbf{r}) = \dfrac{\sum_{i} w_i \cdot c_i}{\sum_{i} w_i},
\quad w_i = \dfrac{1}{\rho_i^2}.
\]

When $\exists i$, where $\rho_i = 0$, then we have to collect all possible $\rho$ values where the distance is also 0, and calculate the mean of their corresponding consequent values as the result.

\section{Rulebases and Surfaces}

The defined rules results a \textit{rule surface} in the higher dimensional space.

% \section{Connection of the Rulebases}

% \section{Evaluation and Usage}

% \section{Example}

\pagebreak

\section{Approximation}

\begin{verbatim}
In universe x
  0 is low and
  1 is high.

In universe y
  0 is low and
  1 is high.

In universe z
  0 is low and
  1 is high.

The z is low when y is low.
The z is low when y is high.
The z is high when x is low.
The z is high when x is high.
\end{verbatim}

There is no $\delta, \rho, \sigma$ operators, which is able to interpolate all of the rule points.

For most of the operators, the set of rules and the rule surface are disjoint.

\end{document}
